The {TestIdeals} package for \emph{Macaulay2}

### 2018-10-16 Fsing/TestIdealsPaper/TestIdeals.tex
Added explanation of HSLG module stuff.

Again, consider the maps $T^e : \omega_{R^{1/p^e}} \to \omega_R$ we have considered throughout this section.  It is a theorem of Hartshorne-Speiser, Lyubeznik, and Gabber \cite{HartshorneSpeiserLocalCohomologyInCharacteristicP,LyubeznikFModulesApplicationsToLocalCohomology,Gabber.tStruc} that the descending images
\begin{equation}\label{eq: descending images}
\omega_R \supseteq T(\omega_{R^{1/p}}) \supseteq \dots \supseteq T^e(\omega_{R^{1/p^e}}) \supseteq T^{e+1}(\omega_{R^{1/p^{e+1}}}) \supseteq \cdots
\end{equation}
stabilize for $e \gg 0$.
The function \texttt{HSLGModule} computes the stable submodule in this chain, called the \emph{HSLG module}, and returns a list containing the following items:
\begin{enumerate}[(1)]
   \item The HSLG module;
   \item The canonical module;
   \item The element $u$ representing the map on the canonical module (see
   \autoref{subsec.DualToFrobeniusOnQuotientRings});
   \item The value of $e > 0$ at which the descending images in \eqref{eq: descending images} stabilize.
\end{enumerate}

@@

More generally, for any ideal $J$ with a map $\phi : J^{1/p^e} \to J$, we have that the images
\[
J \supseteq \phi(J^{1/p^e}) \supseteq \phi^2(J^{1/p^{2e}}) \supseteq \cdots
\]
stabilize as well (in fact, the analogous result even holds for modules).

>>>>

Again, consider the maps $T^e : \omega_{R^{1/p^e}} \to \omega_R$ we have considered throughout this section.  It is a theorem of Hartshorne-Speiser, Lyubeznik, and Gabber \cite{HartshorneSpeiserLocalCohomologyInCharacteristicP,LyubeznikFModulesApplicationsToLocalCohomology,Gabber.tStruc} that the descending images
\begin{equation}\label{eq: descending images}
\omega_R \supseteq T(\omega_{R^{1/p}}) \supseteq \dots \supseteq T^e(\omega_{R^{1/p^e}}) \supseteq T^{e+1}(\omega_{R^{1/p^{e+1}}}) \supseteq \cdots
\end{equation}
stabilize for $e \gg 0$.
The function \texttt{HSLGModule} computes the stable submodule in this chain, called the \emph{HSLG module}, and returns a list containing the following items:
\begin{enumerate}[(1)]
   \item The HSLG module;
   \item The canonical module;
   \item The element $u$ representing the map on the canonical module (see
   \autoref{subsec.DualToFrobeniusOnQuotientRings});
   \item The value of $e > 0$ at which the descending images in \eqref{eq: descending images} stabilize.  This is sometimes also called the HSLG-number of the canonical module as a Cartier module.
\end{enumerate}

@@

More generally, for any ideal $J$ with a map $\phi : J^{1/p^e} \to J$, we have that the images
\[
J \supseteq \phi(J^{1/p^e}) \supseteq \phi^2(J^{1/p^{2e}}) \supseteq \cdots
\]
stabilize as well (in fact, the analogous result even holds for modules \cite{Gabber.tStruc}).

\subsection{Computing the level of a polynomial}
Another interesting invariant that can be calculated using this package
is the so--called \emph{level} of a polynomial; more precisely:

\begin{definition}
Let $K$ be an $F$--finite field of prime characteristic $p,$ and let
$f\in R=K[x_1,\ldots, x_d].$ We define the \emph{level of} $f$ as the
smallest possible integer $e$ where the descending chain
\[
R=(f^{p^0-1})^{[1/p^0]}\supseteq (f^{p-1})^{[1/p]}\supseteq
(f^{p^2-1})^{[1/p^2]}\supseteq\ldots%\supseteq (f^{p^{i-1}-1})^{[1/p^{i-1}]}
\supseteq (f^{p^i-1})^{[1/p^i]}\supseteq\ldots
\]
stabilizes.
\end{definition}
This invariant was introduced in \cite{AlvarezBlickleLyubeznik2005}.  It is also essentially the same data as the HSLG-number of the pair $(R, f^1)$ as computed in (4) above (it is that number $+1$).
One interesting particular
case is when $K=\mathbb{F}_p$ and $f$ is the defining homogeneous polynomial of a hyperelliptic curve
of genus $g;$ when $g=1,$ it was proved in
\cite{BoixDeStefaniVanzo2015} that the corresponding elliptic curve defined by $f$
is ordinary if and only if its level is $1,$ (equivalently, if and only if
$(R/(f))_{(x,y,z)}$ is $F$--pure% \cite{TakTak08}
) and supersingular if and only if its level is $2.$ When
the genus is at least $2,$ level $2$ is a necessary (but not sufficient) condition for the curve
for being ordinary \cite{BlancoBoixFordhamYilmaz2018}; in this case, one also
has that, if the curve is supersingular, then its level has to be at least $3,$ so the level can
always distinguish these two properties in any genus. We illustrate these results by means
of the following examples.

\medskip
{\small
\begin{MyVerbatim}
i1 : R = ZZ/2[x,y,z];
i2 : f = x^3+y^2*z+y*z^2;
i3 : frobeniusPower( 1/2, ideal f )
o3 = ideal (x, y , z )
o3 : Ideal of R
i4 : u=f^3;
i5 : frobeniusPower( 1/4, ideal u )
o6 = ideal (x, y , z )
o6 : Ideal of R
i7 : ((HSLGModule(1, f))#3) + 1
o7 : 2
\end{MyVerbatim}
}\medskip

This shows that this elliptic curve has level $2,$ hence it is supersingular. The next example
shows a non--ordinary hyperelliptic curve of genus $2$ with level $2.$

\medskip
{\small
\begin{MyVerbatim}
i1 : R = ZZ/11[x,y,z];
i2 : f = y^2*z^3-x^5-2*z^5;
i3 : frobeniusPower( 1/11, ideal f )
             2        3
o3 = ideal (z , x*z, x )
o3 : Ideal of R
i4 : u=f^{121-1};
i5 : frobeniusPower( 1/121, ideal u )
             2        3
o6 = ideal (z , x*z, x )
o6 : Ideal of R
i7 : ((HSLGModule(1, f))#3) + 1
o7 : 2
\end{MyVerbatim}
}\medskip
